---
title: "Historical gas budget burn-down"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)

load(here::here("data", "cumulative_heat_models.Rdata"))

QUOTA_M3 <- 1729

# Per-year gas consumption (Spent resets ~0 each Jan)
# Drop years with <100 obs or obvious data errors (>5000 m3)
dat_years <- obs_days_complete %>%
  filter(Spent > 0) %>%
  group_by(year) %>%
  mutate(
    gas_this_year = Spent - min(Spent, na.rm = TRUE),
    gas_remaining = QUOTA_M3 - gas_this_year,
    plot_date     = make_date(year, 1, 1) +
      days(day_in_year - 1)
  ) %>%
  ungroup()

# Filter to usable years
year_stats <- dat_years %>%
  group_by(year) %>%
  summarise(
    n_obs       = n(),
    max_gas     = max(gas_this_year, na.rm = TRUE),
    .groups     = "drop"
  ) %>%
  filter(n_obs >= 100, max_gas < 5000)

dat_years <- dat_years %>%
  filter(year %in% year_stats$year)
```

# Calendar burn-down by year

Each panel shows gas budget remaining (1729 m³ − consumed) over the calendar year. The red dashed line marks the quota boundary — dropping below zero means exceeding the subsidized limit.

```{r fig-burndown-facet}
#| fig-cap: "Gas budget burn-down by year"
#| fig-height: 8
#| fig-width: 10

dat_years %>%
  ggplot(aes(x = plot_date, y = gas_remaining)) +
  theme_bw() +
  geom_line(color = "navy", linewidth = 0.8) +
  geom_hline(yintercept = 0, color = "red",
             linetype = "dashed") +
  facet_wrap(~ year, scales = "free_x", ncol = 2) +
  scale_x_date(date_breaks = "3 months",
               date_labels = "%b") +
  labs(
    x     = "",
    y     = "Gas budget remaining (m\u00b3)",
    title = "Annual gas budget burn-down"
  )
```

# Heat need vs. gas consumed

Alternative parametrisation: x-axis is cumulative heating need (°C·days), y-axis is gas consumed (m³). Each year traces a curve from the origin. The diagonal represents the efficiency ratio — steeper means more gas per degree-day. The black box marks the annual quota (1729 m³) against the typical annual heat need (~3300 °C·days).

```{r fig-heat-vs-gas-facet}
#| fig-cap: "Gas consumed as a function of cumulative heat need"
#| fig-height: 8
#| fig-width: 10

# Heat need also needs to be per-year
dat_heat_gas <- dat_years %>%
  filter(tavg_low_cumul > 0, gas_this_year > 0)

dat_heat_gas %>%
  ggplot(aes(x = tavg_low_cumul, y = gas_this_year,
             color = factor(year))) +
  theme_bw() +
  geom_line(linewidth = 0.8) +
  # Quota ceiling
  geom_hline(yintercept = QUOTA_M3, color = "black",
             linetype = "dashed", linewidth = 0.8) +
  # Typical end-of-year heat need marker
  geom_vline(xintercept = 3300, color = "grey50",
             linetype = "dotted") +
  labs(
    x     = "Cumulative heating need (\u00b0C\u00b7days)",
    y     = "Gas consumed (m\u00b3)",
    color = "Year",
    title = "Gas consumption vs. cumulative heat need"
  )
```

Same data faceted by year for clarity:

```{r fig-heat-vs-gas-individual}
#| fig-cap: "Gas consumed vs. heat need (per year)"
#| fig-height: 8
#| fig-width: 10

dat_heat_gas %>%
  ggplot(aes(x = tavg_low_cumul, y = gas_this_year)) +
  theme_bw() +
  geom_line(color = "darkgreen", linewidth = 0.8) +
  geom_hline(yintercept = QUOTA_M3, color = "red",
             linetype = "dashed") +
  # Ideal efficiency slope (quota / typical annual heat)
  geom_abline(
    slope     = QUOTA_M3 / 3300,
    intercept = 0,
    color     = "salmon", linetype = "dotted",
    linewidth = 0.8
  ) +
  facet_wrap(~ year, ncol = 2) +
  labs(
    x     = "Cumulative heating need (\u00b0C\u00b7days)",
    y     = "Gas consumed (m\u00b3)",
    title = "Heat need vs. gas: ideal slope (dotted) vs. actual"
  )
```

The dotted salmon line in each panel represents the "ideal" efficiency: consuming exactly 1729 m³ for ~3300 °C·days of annual heat. Years where the green line runs above the salmon line are burning gas faster than the budget allows; below means headroom.
