---
title: "child2"
format: docx
editor: visual
---

# Quantifying cold and long winters

The amount of energy required for heating is the function of the outside temperature.

Therefore to predict the energy needs left in a given (heating) season, we should make a prediction about the expected amount of energy required until the end of the season.

Thankfully things seem to be straightforward if we assume that the inside temperature of the building should be constant (it's not but bear with me). Assume 20°C (68°F). Heat loss is linear to the difference in temperature between the inside and outside temperatures [link](http://hyperphysics.phy-astr.gsu.edu/hbase/thermo/heatloss.html).

Therefore when the outside temperature is 16°C we would need 4x the power to heat the building than if it would be 19°C. Also, this amount of energy needed for these two days is the same that we would need for heating for 3 days if it were $20 - 5/3 = 18.33$ degrees outside. In other words for a given heating season, we would add up the "missing degrees" until 20 for every day the outside temperature was less than 20°.

```{r fig-heatneed-per-year}
#| fig-cap: "Cumulative heating energy needs by year"

# This figure is produced by iter4_cumulative_heat_model.r
# obs_days_complete is loaded via cumulative_heat_models.Rdata
obs_days_complete %>%
  ggplot(aes(x = day_in_wint, y = tavg_low_cumul)) +
  theme_bw() +
  geom_line(alpha = .7,
            mapping = aes(color = year,
                          fill = year,
                          group = factor(ywint))) +
  labs(x = "Day in the season (starts Aug.1st)",
       y = "Heating energy needs [°C*Day]")
```

I personally like this approach very much, but of course we are neglecting:

-   The effects of solar radiation, eg. sunlight heats up the building a bit,
-   the building's heat capacity, eg. after a hot summer's day, it takes about 3 days for the walls to cool

For simplicity and to retain my sanity, lets also assume that at each timepoint (day), the distribution of 'energy needs' is normal.

I'd like to fit a *linear model* for the energy needs with a natural spline term containing the elapsed days. To capture the apparent heteroscedasticity, I'd opt for a *generalized least squares* model with an *exponential variance structure* (meaning that the residuals are understood to have a larger variance when more days have elapsed).

```{r fig-gls-effects}
#| fig-cap: "GLS model effects"

# mod_cum and sds are from cumulative_heat_models.Rdata
mod_cum %>%
  effects::predictorEffects(partial.residuals = FALSE) %>%
  plot()
```

As a "petty craftsman" kind of approach, I calculated standard deviations for all days and then fitted a *linear model* over it with a natural spline term capturing the evolution of standard deviation in heating needs over the course of a season.

```{r fig-sd-effects}
#| fig-cap: "Observed SD evolution over the season"

mod_cum_obs %>%
  effects::predictorEffects(partial.residuals = TRUE) %>%
  plot()
```

Now we have a mean and a variance for the cumulative heating energy needs on any given day (at my location)! I can then assign a z-score for each result through the years.

```{r fig-cumulative-with-bands}
#| fig-cap: "Cumulative heating needs with mean and 95% CI bands"

obs_days_complete %>%
  ggplot(aes(x = day_in_wint, y = tavg_low_cumul)) +
  theme_bw() +
  geom_line(alpha = .7,
            mapping = aes(color = year,
                          fill = year,
                          group = factor(ywint))) +
  geom_line(data = sds,
            mapping = aes(y = mean_pred),
            color = "salmon4", linewidth = 1.5) +
  geom_line(data = sds,
            mapping = aes(y = mean_pred + 1.96 * sd_pred),
            color = "red", linewidth = 1.5) +
  geom_line(data = sds,
            mapping = aes(y = mean_pred - 1.96 * sd_pred),
            color = "red", linewidth = 1.5)
```


# Estimating total energy needs left on a given day

This is all well and good, and this translates to an estimated total energy need at the beginning of a season, but we are mainly interested in the *conditional* energy need for the total season if we have observed the cumulative energy need up to day "*d".*

I start with investigating how do energy need values observed on a given day *correlate* to the *total* energy need for a given season. Specifically, what was the correlation between the z-scores of a given day vs. the last day in a season.

```{r fig-correlation-evolution}
#| fig-cap: "Correlation between mid-season z-score and end-of-season z-score"

sds %>%
  ggplot(aes(x = day_in_wint, y = cor_z)) +
  theme_bw() +
  geom_line()
```

The spike around day 150 is due to leap years (probably); didn't bother to fix it, as it didn't bother the results much.


# Figuring in actual energy usage

The key insight is that a linear budget makes no sense — winter months consume vastly more gas than summer. Instead we track the **ratio of gas consumed to cumulative heating need** (m³ per °C·day). If this ratio stays stable, we can predict total annual gas usage by combining it with the *conditional* total heat need forecast.

```{r fig-spent-vs-heatneed}
#| fig-cap: "Gas consumed vs. cumulative heat need (all years)"

# Gas consumed (m3) plotted against cumulative heat need
# Black box = ideal budget (1729 m3 for ~3284 °C·days)
# Salmon slope = mean efficiency across years
QUOTA_M3 <- 1729

# Only keep seasons that start near (0,0): the first
# observation in a season should have low Spent AND low
# tavg_low_cumul. Drop seasons with a broken origin.
dat_plot_heatneed <- obs_days_complete %>%
  filter(Spent > 0, tavg_low_cumul > 0) %>%
  group_by(ywint) %>%
  filter(min(tavg_low_cumul, na.rm = TRUE) < 50) %>%
  ungroup()

dat_plot_heatneed %>%
  ggplot(aes(x = tavg_low_cumul, y = Spent,
             color = factor(year(Date)))) +
  geom_point(alpha = 0.6, size = 0.8) +
  geom_smooth(method = "lm",
              mapping = aes(color = "smootho"),
              formula = y ~ x - 1,
              color = "grey50") +
  # Full-year budget box (1729 m3 / ~3284 deg-days)
  geom_line(data = data.frame(
    x = c(0, 3284), y = c(QUOTA_M3, QUOTA_M3)),
    mapping = aes(x, y),
    color = "black", linewidth = 1) +
  geom_line(data = data.frame(
    x = c(3284, 3284), y = c(QUOTA_M3, 0)),
    mapping = aes(x, y),
    color = "black", linewidth = 1) +
  geom_line(data = data.frame(
    x = c(3284, 0), y = c(QUOTA_M3, 0)),
    mapping = aes(x, y),
    color = "black", linewidth = 1) +
  # Mean efficiency slope (salmon)
  geom_line(data = data.frame(
    x = c(0, 3000), y = c(0, 2000)),
    mapping = aes(x, y),
    color = "salmon4", linewidth = 2) +
  theme_minimal() +
  labs(x = "Cumulative heat need (°C·days)",
       y = "Gas consumed (m³)",
       color = "Year")
```

## Seasonal efficiency tracking

Rather than a flat daily budget, we track the *gas consumption efficiency* — how many m³ of gas per °C·day of heating need. This ratio should stay roughly stable over a season. If it climbs, we are over-heating; if it drops, we are under-heating (or the heat pump is picking up slack).

The dashed grey lines show the predicted efficiency band for the remaining season, based on conditional forecasts of total heat need.

```{r fig-efficiency-dashboard}
#| fig-cap: "Gas efficiency ratio: current year vs. predicted band"

# Determine the current/latest ywint
act_ywint <- obs_days_complete %>%
  ungroup() %>%
  filter(Date == max(Date, na.rm = TRUE)) %>%
  slice(1) %>%
  pull(ywint)

# X-axis: Jan 1 of the current year to Jul 31
# (practical view — meter was read ~Jan 5, this is
# the window we care about for budget tracking)
current_year <- lubridate::year(Sys.Date())
date_limits <- c(
  as.POSIXct(paste0(current_year, "-01-01")),
  as.POSIXct(paste0(current_year, "-07-31"))
)

# The key plot: efficiency ratio over the season
# Color gradient: blue (lots of gas left) -> green (tight)
obs_days_complete %>%
  filter(ywint == act_ywint,
         Date >= date_limits[1]) %>%
  ggplot(aes(x = Date, y = spent_tavg,
             color = gas_left)) +
  theme_bw() +
  scale_color_gradient(
    high = "blue", low = "green",
    limits = c(0, QUOTA_M3),
    name = "Gas left (m³)") +
  geom_line(linewidth = 1.5) +
  # Predicted mean efficiency for remaining season
  geom_line(aes(y = spent_tavg_left_mean),
            color = "grey50",
            linetype = "dashed", linewidth = 1.2) +
  # 95% prediction band
  geom_line(aes(y = spent_tavg_left_lwr),
            color = "grey70",
            linetype = "dashed", linewidth = 1.2) +
  geom_line(aes(y = spent_tavg_left_upr),
            color = "grey70",
            linetype = "dashed", linewidth = 1.2) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_datetime(
    date_breaks = "1 month",
    date_labels = "%b %Y",
    limits = date_limits) +
  labs(x = "", y = "Gas consumption (m³ per °C·day)",
       title = "Seasonal efficiency: Jan–Jul view")
```

The solid colored line shows actual gas usage efficiency (m³ per °C·day of heat need), colored by remaining gas budget. When the line stays within the grey dashed prediction bands, the budget is on track. If the line moves above the upper band, gas consumption is too aggressive relative to expected remaining heat need.
