---
title: "child2"
format: html
editor: visual
---

# Quantifying cold and long winters

The amount of energy required for heating is a function of the outside temperature.

Therefore to predict the energy needs left in a given year, we should make a prediction about the expected amount of energy required until December 31.

Assume a constant indoor temperature of 20 °C. Heat loss is linear to the inside-outside temperature difference. Therefore when the outside temperature is 16 °C we need 4x the power than if it were 19 °C. For a given year we add up the "missing degrees" below 20 for every day — the **cumulative heating need** in °C·days.

```{r fig-heatneed-per-year}
#| fig-cap: "Cumulative heating energy needs by calendar year"

# obs_days_complete is loaded via cumulative_heat_models.Rdata
obs_days_complete %>%
  ggplot(aes(x = day_in_year, y = tavg_low_cumul)) +
  theme_bw() +
  geom_line(alpha = .7,
            mapping = aes(color = year,
                          fill = year,
                          group = factor(year))) +
  labs(x = "Day of year",
       y = "Heating energy needs [°C*Day]")
```

We are neglecting solar radiation and building heat capacity. For simplicity, assume that at each day the distribution of cumulative heating needs across years is approximately normal.

A **GLS model** with a natural spline (8 df) captures the seasonal curve. An exponential variance structure accounts for increasing spread as the year progresses.

```{r fig-gls-effects}
#| fig-cap: "GLS model effects"

# mod_cum and sds are from cumulative_heat_models.Rdata
mod_cum %>%
  effects::predictorEffects(partial.residuals = FALSE) %>%
  plot()
```

Standard deviations for each day of year, smoothed with a natural spline:

```{r fig-sd-effects}
#| fig-cap: "Observed SD evolution over the year"

mod_cum_obs %>%
  effects::predictorEffects(partial.residuals = TRUE) %>%
  plot()
```

Now we have a mean and variance for cumulative heating needs on any day. We assign z-scores to every observed year.

```{r fig-cumulative-with-bands}
#| fig-cap: "Cumulative heating needs with mean and 95% CI bands"

obs_days_complete %>%
  ggplot(aes(x = day_in_year, y = tavg_low_cumul)) +
  theme_bw() +
  geom_line(alpha = .7,
            mapping = aes(color = year,
                          fill = year,
                          group = factor(year))) +
  geom_line(data = sds,
            mapping = aes(y = mean_pred),
            color = "salmon4", linewidth = 1.5) +
  geom_line(data = sds,
            mapping = aes(y = mean_pred + 1.96 * sd_pred),
            color = "red", linewidth = 1.5) +
  geom_line(data = sds,
            mapping = aes(y = mean_pred - 1.96 * sd_pred),
            color = "red", linewidth = 1.5)
```


# Estimating total energy needs left on a given day

We are mainly interested in the *conditional* energy need for the total year given what we have observed up to day *d*.

How well does the z-score on a given day predict the end-of-year z-score?

```{r fig-correlation-evolution}
#| fig-cap: "Correlation between current z-score and end-of-year z-score"

sds %>%
  ggplot(aes(x = day_in_year, y = cor_z)) +
  theme_bw() +
  geom_line()
```


# Figuring in actual energy usage

We track the **ratio of gas consumed to cumulative heating need** (m³ per °C·day). If this ratio stays stable, we can predict total annual gas usage by combining it with the *conditional* total heat need forecast.

```{r fig-spent-vs-heatneed}
#| fig-cap: "Gas consumed vs. cumulative heat need (all years)"

QUOTA_M3 <- 1729

# Only keep years that start near (0,0)
dat_plot_heatneed <- obs_days_complete %>%
  filter(Spent > 0, tavg_low_cumul > 0) %>%
  group_by(year) %>%
  filter(min(tavg_low_cumul, na.rm = TRUE) < 50) %>%
  ungroup()

dat_plot_heatneed %>%
  ggplot(aes(x = tavg_low_cumul, y = Spent,
             color = factor(year))) +
  geom_point(alpha = 0.6, size = 0.8) +
  geom_smooth(method = "lm",
              mapping = aes(color = "smootho"),
              formula = y ~ x - 1,
              color = "grey50") +
  # Full-year budget box (1729 m3 / ~3284 deg-days)
  geom_line(data = data.frame(
    x = c(0, 3284), y = c(QUOTA_M3, QUOTA_M3)),
    mapping = aes(x, y),
    color = "black", linewidth = 1) +
  geom_line(data = data.frame(
    x = c(3284, 3284), y = c(QUOTA_M3, 0)),
    mapping = aes(x, y),
    color = "black", linewidth = 1) +
  geom_line(data = data.frame(
    x = c(3284, 0), y = c(QUOTA_M3, 0)),
    mapping = aes(x, y),
    color = "black", linewidth = 1) +
  # Mean efficiency slope (salmon)
  geom_line(data = data.frame(
    x = c(0, 3000), y = c(0, 2000)),
    mapping = aes(x, y),
    color = "salmon4", linewidth = 2) +
  theme_minimal() +
  labs(x = "Cumulative heat need (°C·days)",
       y = "Gas consumed (m³)",
       color = "Year")
```

## Efficiency tracking (Jan–Dec)

Rather than a flat daily budget, we track the *gas consumption efficiency* — how many m³ of gas per °C·day of heating need. This ratio should stay roughly stable. If it climbs, we are over-heating; if it drops, the heat pump is picking up slack.

The dashed grey lines show the predicted efficiency band based on conditional forecasts of total heat need.

```{r fig-efficiency-dashboard}
#| fig-cap: "Gas efficiency ratio: current year vs. predicted band"

# X-axis: Jan 1 to Dec 31 of current year
current_year <- lubridate::year(Sys.Date())
date_limits <- c(
  as.POSIXct(paste0(current_year, "-01-01")),
  as.POSIXct(paste0(current_year, "-12-31"))
)

# The key plot: efficiency ratio over the year
# Color gradient: blue (lots of gas left) -> green (tight)
obs_days_complete %>%
  filter(year == current_year) %>%
  ggplot(aes(x = Date, y = spent_tavg,
             color = gas_left_jan)) +
  theme_bw() +
  scale_color_gradient(
    high = "blue", low = "green",
    limits = c(0, 1729),
    name = "Gas left (m³)") +
  geom_line(linewidth = 1.5) +
  # Predicted mean efficiency for remaining year
  geom_line(aes(y = spent_tavg_left_mean),
            color = "grey50",
            linetype = "dashed", linewidth = 1.2) +
  # 95% prediction band
  geom_line(aes(y = spent_tavg_left_lwr),
            color = "grey70",
            linetype = "dashed", linewidth = 1.2) +
  geom_line(aes(y = spent_tavg_left_upr),
            color = "grey70",
            linetype = "dashed", linewidth = 1.2) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_datetime(
    date_breaks = "1 month",
    date_labels = "%b",
    limits = date_limits) +
  labs(x = "", y = "Gas consumption (m³ per °C·day)",
       title = "Efficiency tracking: Jan–Dec")
```

The solid colored line shows actual gas usage efficiency, colored by remaining gas budget. When the line stays within the grey dashed prediction bands, the budget is on track.
