---
title: "Gas budget dashboard"
subtitle: "Should I amp up, hold steady, or dial down this week?"
author: "Márton Kiss, MD"
date: today
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    toc-title: "Jump to"
    page-layout: full
    fig-width: 9
    fig-height: 5.2
    code-fold: true
    embed-resources: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Load packages, functions, and core data
source(here::here("inst", "function", "load_stuff.r"))

# Analysis outputs (produced by the iter1-iter5 pipeline)
load(here::here("data", "cumulative_heat_models.Rdata"))
load(here::here("data", "cumulative_heat_predictions.Rdata"))
load(here::here("data", "decision_snapshot.Rdata"))

ds <- decision_snapshot

# Helper: round + format a number with a unit
fmt <- function(x, unit = "", digits = 0) {
  paste0(formatC(round(x, digits), format = "f",
                 big.mark = " ", digits = digits), unit)
}

# Colour-code the verdict
verdict_col <- switch(
  ds$decision,
  "AMP UP"      = "#2e7d32",  # green: room to spare
  "HOLD STEADY" = "#f9a825",  # amber: tight
  "DIAL DOWN"   = "#c62828",  # red: at risk
  "#555555"
)
```

```{r verdict-banner, results = "asis"}
# A prominent, plain-language verdict banner. This is the
# single most important takeaway in the whole report.
cat(sprintf('
<div style="border-left: 10px solid %s; background: %s14;
            padding: 18px 22px; border-radius: 6px;
            margin: 8px 0 24px 0;">
  <div style="font-size: 0.85em; letter-spacing: 1px;
              text-transform: uppercase; color: #777;">
    This week&rsquo;s recommendation
  </div>
  <div style="font-size: 2.1em; font-weight: 700;
              color: %s; line-height: 1.1; margin: 4px 0;">
    %s
  </div>
  <div style="font-size: 1.05em; color: #333;">%s</div>
</div>',
  verdict_col, verdict_col, verdict_col,
  ds$decision, ds$decision_detail))
```

## The burn-down chart

This is the heart of the report. It answers one question:
**will the year&rsquo;s subsidised gas allowance (1 729 m³) last
until December 31?**

```{r burndown, fig.height = 5.4}
fig_burndown
```

::: {.callout-tip collapse="false"}
## How to read this chart

- **Solid navy line** &mdash; gas budget actually remaining,
  reconstructed from your meter readings. This part is
  *certain*: it already happened.
- **Dashed red line** &mdash; the *expected* trajectory for the
  rest of the year, based on 30 years of weather and your
  current consumption efficiency.
- **Shaded pink band** &mdash; the 95 % range of outcomes. It is
  narrow now and **fans out toward winter**, because the single
  biggest unknown is *how cold the coming heating season will
  be*. A mild autumn keeps you near the top of the band; a harsh
  one pushes you toward the bottom.
- **Red horizontal line (0 m³)** &mdash; the quota boundary.
  Anything below it is gas billed at the **penalty price**
  (~7.5&times; the subsidised rate).
:::

## Where the budget stands today

```{r status-table}
library(knitr)

status_tbl <- data.frame(
  Metric = c(
    "Tracking window",
    "Days elapsed",
    "Gas consumed since Jan 1",
    "Gas remaining in quota",
    "Expected need for rest of year (95% CI)",
    "Projected year-end balance (95% CI)",
    "Chance of overshooting the quota",
    "Current daily rate (latest meter interval)",
    "Daily rate to stay on budget"
  ),
  Value = c(
    sprintf("Jan 1 &ndash; Dec 31, %d", ds$current_year),
    sprintf("%d of %d", round(ds$days_elapsed),
            round(ds$days_elapsed + ds$days_remaining)),
    fmt(ds$gas_consumed, " m³"),
    fmt(ds$gas_budget_left, " m³"),
    sprintf("%s &ndash; %s",
            fmt(ds$gas_need["lower"], " m³"),
            fmt(ds$gas_need["upper"], " m³")),
    sprintf("%s (%s to %s)",
            fmt(ds$forecast_year_end["mean"], " m³"),
            fmt(ds$forecast_year_end["lower"], " m³"),
            fmt(ds$forecast_year_end["upper"], " m³")),
    sprintf("~%d%%", round(ds$prob_overshoot)),
    fmt(ds$current_daily_rate, " m³/day", 2),
    sprintf("%s &ndash; %s",
            fmt(ds$daily_rate_suggested["lower"], "", 2),
            fmt(ds$daily_rate_suggested["upper"], " m³/day", 2))
  ),
  check.names = FALSE
)

kable(status_tbl, escape = FALSE, align = c("l", "r"))
```

```{r hp-note, results = "asis"}
# Heat-pump hint when it is warm enough for the AC to win
if (!is.na(ds$recent_tavg) && ds$recent_tavg > 10) {
  cat(sprintf('
::: {.callout-note}
Recent 3-day average outdoor temperature is **%.1f&nbsp;°C**
(above 10&nbsp;°C). The heat pump is the cheaper way to add
heat right now &mdash; save the discounted gas for the cold
months.
:::', ds$recent_tavg))
}
```

# Supporting detail {#supporting}

The verdict above is driven by two ideas: (1) how much *heating
need* is still ahead of us, and (2) how much *gas* we burn per
unit of heating need. The next two charts show each.

## Efficiency: gas burned per unit of heating need

We track gas used per °C·day of heating need. If this ratio
holds steady, we can convert a weather forecast into a gas
forecast. The grey band is the predicted efficiency corridor
for the remainder of the year.

```{r fig-efficiency}
fig_efficiency_ratio
```

## How predictable is the rest of the year?

Early in the year, the final heating total is highly uncertain.
As the season unfolds, today&rsquo;s running total becomes a
better and better predictor of where we will end up &mdash;
which is why the burn-down band narrows as winter approaches.

```{r fig-bands}
fig_conditional_bands
```

# Methodology {#methodology}

::: {.callout-note collapse="true"}
## Modelling notes (click to expand)

**Heating need.** Assuming a constant indoor target of 20 °C and
heat loss proportional to the indoor&ndash;outdoor temperature
gap, we sum the "missing degrees" below 20 °C each day to get a
**cumulative heating need** in °C·days.

**Weather model.** A GLS model with a natural spline (8 df)
captures the seasonal mean of cumulative heating need across
~30 years of Budapest weather (station 12843). An exponential
variance structure lets the spread grow through the year.

**Conditional forecast.** Given the heating need observed *so
far this year*, the end-of-year total is forecast with a
conditional-normal model: the correlation between today&rsquo;s
running total and each future day (estimated empirically across
years) shrinks the uncertainty band the further into the season
we are.

**Gas from heat.** Gas use is split into two parts: a per-day
**hot-water baseline** (present even when no heating is needed)
and **heating gas** proportional to the cumulative heating
need. The baseline is estimated from this year&rsquo;s data as
the marginal gas-per-day controlling for heat need; when that
estimate is unreliable (collinear within a single winter, so
it can even go negative) it falls back to a dedicated
0.4&nbsp;m³/day. The same baseline and heating efficiency feed
both the burn-down and the efficiency chart.

**Budget window.** Tracking runs on the **calendar year
(Jan 1 &ndash; Dec 31)**, anchored to the meter value
interpolated at Jan 1. (The official subsidised-quota year is
Aug&ndash;Jul, but the practical átalány true-up happens in
early January, so the calendar year is the relevant window
here.)

**Pipeline.** `iter1` (data prep) → `iter2` (consumption
curves) → `iter3` (cumulative heat model) → `iter4` (conditional
prediction) → `iter5` (decision tool). Re-run after each new
meter reading.
:::

```{r compile-time, results = "asis"}
cat("\n\n*Compiled ",
    as.character(Sys.time()), ".*", sep = "")
```
